---
title: "titanic_competition"
author: "Salma Elshahawy"
date: "12/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(dplyr)
library(caret)
library(readr)
```

```{r}
train_df <- read.csv("train.csv", stringsAsFactors = FALSE)
train_df$Survived <- as.factor(train_df$Survived)
head(train_df)
paste("Mean is: ",mean(train_df$Age))
paste("Median is: " ,median(train_df$Age))
paste("Mode is: " ,mode(train_df$Age))
paste("Standard deviation is: " ,sd(train_df$Age))
```

```{r}
# Impute missing ages in Training data

train_df %>% select(-Ticket, -Name, -Cabin, -Embarked) -> train_ages
pre_proc <- preProcess(train_ages, method = "bagImpute")
train_ages <- predict(pre_proc, train_ages)
train_df$Age <- train_ages$Age
head(train_df)
```
```{r}
prop.table(table(train_df$Sex,train_df$Survived),margin=1)
```


```{r}
train_df %>%
  mutate(family_size = SibSp + Parch + 1) -> train_df

train_df
```

```{r}
# plot sibsp vs parch - survival/dead cat/cat
p4 <- ggplot(train_df, aes(factor(family_size), fill = Survived)) + 
  geom_bar() +
  facet_wrap(~Pclass*Sex)

p4
```

```{r}
ggplot(train_df, mapping = aes(x = Survived, y = Age, fill = Survived)) + 
  geom_boxplot(outlier.colour = "red", outlier.shape = 5, outlier.size = 4) + 
  facet_wrap(~Sex)
paste("Mean is: ",mean(train_df$Age))
paste("Median is: " ,median(train_df$Age))
paste("Mode is: " ,mode(train_df$Age))
paste("Standard deviation is: " ,sd(train_df$Age))
```

We had some outliers here need to be normalized for the age. 

# Outlier in Age imputation 

```{r}
capOutlier <- function(x){
   qnt <- quantile(x, probs=c(.25, .75))
   caps <- quantile(x, probs=c(.05, .95))
   H <- 1.5 * IQR(x, na.rm = T)
   x[x < (qnt[1] - H)] <- caps[1]
   x[x > (qnt[2] + H)] <- caps[2]
   return(x)
}
train_df$Age=capOutlier(train_df$Age)
ggplot(train_df, mapping = aes(x = Survived, y = Age, fill = Survived)) + 
  geom_boxplot(outlier.colour = "red", outlier.shape = 8, outlier.size = 4) + 
  facet_wrap(~Sex)
paste("Mean is: ",mean(train_df$Age))
paste("Median is: " ,median(train_df$Age))
paste("Mode is: " ,mode(train_df$Age))
paste("Standard deviation is: " ,sd(train_df$Age))
```


```{r}
ggplot(train_df, aes(factor(Embarked), fill = Survived)) + 
  geom_bar() + 
  facet_wrap(~Pclass)
# Southampton seems to be the most embarked that had survivals
```


Do some feature engineering by extracting extra embeded information from our data
First, Extract title from name as a new column

```{r}
# train_df$Title <- sapply(train_df$Name, FUN=function(x) {strsplit(x, split='[,.]')[[1]][2]})
train_df %>%
  rowwise %>%
  mutate(title = gsub('(.*, )|(\\..*)', '', Name))-> train_df
train_df$Surname <- sapply(train_df$Name, FUN=function(x) {strsplit(x, split='[,.]')[[1]][1]})

train_df$title[train_df$title %in% c('Mme', 'Mlle')] <- 'Mlle'
train_df$title[train_df$title %in% c('Capt', 'Don', 'Major', 'Sir')] <- 'Sir'
train_df$title[train_df$title %in% c('Dona', 'Lady', 'the Countess', 'Jonkheer')] <- 'Lady'
```

```{r}
library(forcats)
ggplot(train_df, aes(fct_infreq(factor(title)), fill = Survived)) +
  geom_bar() +
  coord_flip() +
  facet_wrap(~Sex)
  ggtitle("Survival by titles")
```


Survived females are more than males - so we are on the right track

Extract ticket number - the first three digit after subressing letters

```{r}
train_df %>%
  mutate(ticket = gsub("([^\\d]+[2,5].\\s)|([A-Z]+\\/[5])|([^\\d])", "", Ticket, perl = TRUE))  -> train_df

train_df$ticket <- substr(train_df$ticket, 0, 3)
transform(train_df, ticket = as.numeric(ticket)) -> train_df
train_df
```

We have some NAs, impute them using `mice package`

```{r}
# clean up NAs
sapply(train_df, function(x) sum(is.na(x)))
# PassengerId    Survived      Pclass        Name         Sex         Age       SibSp       Parch      Ticket        Fare 
#           0           0           0           0           0           0           0           0           0           0 
#       Cabin    Embarked 
#           0           0 
```

```{r}
library(mice)
init = mice(train_df, maxit=0) 
meth = init$method
predM = init$predictorMatrix
predM[, c("PassengerId")]=0
meth[c("ticket")]="norm" 
set.seed(103)
imputed = mice(train_df, method=meth, predictorMatrix=predM, m=5)
train_df <- complete(imputed)
sapply(train_df, function(x) sum(is.na(x)))
```

```{r}
test_df <- read.csv("test.csv", stringsAsFactors = FALSE)
# Impute missing ages in Training data
test_df %>% select(-Ticket, -Name, -Cabin, -Embarked) -> test_ages

pre_proc <- preProcess(test_ages, method = "bagImpute")
test_ages <- predict(pre_proc, test_ages)
test_df$Age <- test_ages$Age
head(test_df)
```

```{r}
test_df %>%
  mutate(family_size = SibSp + Parch + 1) -> test_df

test_df
```

```{r}
capOutlier <- function(x){
   qnt <- quantile(x, probs=c(.25, .75))
   caps <- quantile(x, probs=c(.05, .95))
   H <- 1.5 * IQR(x, na.rm = T)
   x[x < (qnt[1] - H)] <- caps[1]
   x[x > (qnt[2] + H)] <- caps[2]
   return(x)
}
test_df$Age=capOutlier(test_df$Age)
```

```{r}
test_df %>%
  rowwise %>%
  mutate(title = gsub('(.*, )|(\\..*)', '', Name))-> test_df
test_df$Surname <- sapply(test_df$Name, FUN=function(x) {strsplit(x, split='[,.]')[[1]][1]})

test_df$title[test_df$title %in% c('Mme', 'Mlle')] <- 'Mlle'
test_df$title[test_df$title %in% c('Capt', 'Don', 'Major', 'Sir')] <- 'Sir'
test_df$title[test_df$title %in% c('Dona', 'Lady', 'the Countess', 'Jonkheer')] <- 'Lady'
```

```{r}
test_df %>%
  mutate(ticket = gsub("([^\\d]+[2,5].\\s)|([A-Z]+\\/[5])|([^\\d])", "", Ticket, perl = TRUE))  -> test_df

test_df$ticket <- substr(test_df$ticket, 0, 3)
transform(test_df, ticket = as.numeric(ticket)) -> test_df
test_df
```

```{r}
library(mice)
init = mice(test_df, maxit=0) 
meth = init$method
predM = init$predictorMatrix
predM[, c("PassengerId")]=0
meth[c("ticket")]="norm" 
set.seed(103)
imputed = mice(test_df, method=meth, predictorMatrix=predM, m=5)
test_df <- complete(imputed)
```



```{r}
trctrl <- trainControl(## 10-fold CV
                           method = "repeatedcv",
                           number = 10,
                           ## repeated ten times
                           repeats = 10)
```

```{r}
set.seed(825)

fit1 <- train(Survived ~ Age + Pclass + Sex + Embarked + ticket + title + family_size + Fare + Cabin,
                 data = train_df,
                 method = "ranger",
                 trControl = trctrl,
                 ## This last option is actually one
                 ## for gbm() that passes through
                 verbose = FALSE)
fit1
```


```{r}
tgctrl <- expand.grid(
    .mtry = 12,
    .splitrule = "gini",
    .min.node.size = 17)

set.seed(1000)
model.cv <- train(Survived ~ Age + Pclass + Sex + Embarked + ticket + title + family_size + Fare,
               data = train_df,
               trControl = trctrl,
               metric = "Accuracy",
               importance = "impurity",
               tuneGrid = tgctrl,
               num.trees = 1500,
               method = "ranger")
model.cv
```


```{r}
ggplot(varImp(model.cv))
```

```{r}
Survived <- predict(model.cv, test_df)
test_df %>%
    cbind(., Survived) -> test_pred
test_pred %>%
  select(PassengerId, Survived) -> test_final
test_final
```

```{r}
ggplot(test_final, aes(fct_infreq(factor(Survived)))) +
  geom_bar()
```

```{r}
write_csv(test_final, 'test_output.csv')
```

